---
title: "Survival TV series"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library('flexdashboard')
library('here')
library('ggplot2')
library("tidytuesdayR")
library("tidyverse")
library("purrr")
library("DT")
library("shiny")
library("showtext")
```

```{r}
# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

if (!file.exists(here("data", "tuesdata_survivalists.csv"))) {
    tuesdata <- tidytuesdayR::tt_load("2023-01-24")
    survivalists <- tuesdata$survivalists
    loadouts <- tuesdata$loadouts

    # save the files to RDS objects
    write.csv(tuesdata$survivalists, file = here("data", "tuesdata_survivalists.csv"))
    write.csv(tuesdata$loadouts, file = here("data", "tuesdata_loadouts.csv"))
}

survivalists <- read.csv(here("data", "tuesdata_survivalists.csv"))
loadouts <- read.csv(here("data", "tuesdata_loadouts.csv"))

```

```{r}
# filter 'country' contains "United States"
has_united_states <- map_lgl(survivalists$country, ~ str_detect(.x, "United States"))
clean_survival <- survivalists[has_united_states, ]

df <- expand_grid(
  days_lasted = 0: max(clean_survival$days_lasted),
  gender = unique(clean_survival$gender)
) |> 
  left_join(
    clean_survival |> 
      count(days_lasted, gender),
      by = c("days_lasted", "gender")
  ) |> 
  left_join(
    clean_survival |> 
      count(gender, name = "N"),
      by = "gender"
  ) |> 
  group_by(gender) |> 
  mutate(
    n = replace_na(n, 0),
    n_lasted = N-cumsum(n),
    p = n_lasted/N
  ) 


clean_load2 <- loadouts |>
    group_by(name, season) |>
    nest() |>
    mutate(
        load_stats = map(.x = data, .f = ~tibble(
            load_minimum = min(.x$item_number),
            load_maximum = max(.x$item_number),
            load_mean = mean(.x$item_number)
        ))
    )  |>
    select(-data)  |>
    unnest(load_stats)

```

# About

## Column  

### **Purpose of the dashboard**

Do some survival analysis for people with different characteristics. <br>
Based on the data frame of survivalists across all 9 seasons, we would like to explore if males live longer than females? How many loadouts every survivalists has taken?

### **Data original**

This data of the TV series [Alone](https://www.history.com/shows/alone) collected and shared on TidyTuesday, which is a weekly podcast and global community activity brought to you by the R4DS Online Learning Community. [Archive of datasets from the 2023 series of *TidyTuesday* events](https://github.com/rfordatascience/tidytuesday/tree/master/data/2023)  <br>

In the survival TV series 'Alone,' 10 survivalists are dropped in an extremely remote area and must fend for themselves. They aim to last 100 days in the Artic winter, living off the land through their survival skills, endurance, and mental fortitude.

[Data Source for this project](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-24/readme.md)  <br>

Here is a data dictionary for what all the column names mean: [data dictionary](https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-01-24#data-dictionary)

### **Author**

Siyu Zou. The project code could be found in [Github](https://github.com/zsyfirst/Project4)


## Column 

### **A video explaining this dashboard**


# The Data 

## Column {data-width="400"}

### **Data description** 

A data frame of survivalists across all 9 seasons detailing demographics, location, ranking, and days lasted.

In the survival TV series ‘Alone,’ 10 survivalists are dropped in an extremely remote area and must fend for themselves. 

It is a collection of datasets about the TV series in a tidy format. Included in this analysis are 2 datasets

*survivalists* 

*loadouts*

## Column {data-width="600"}

### Data table
```{r}
data <- survivalists[, c("season" ,"age","gender","city","state","result" ,"days_lasted")] %>% 
  arrange(season) %>% 
  head(50) %>% 
  setNames(c("Season","Age","Gender","City","State","Rank","Days lasted")) 

renderDataTable({
  datatable(data)})
```

# Survival curve

## Column {.tabset}

### Survival curves

```{r}
plotOutput("myPlot", width = "100%", height = "100%")
df |> 
      ggplot(aes(days_lasted, p, colour = gender)) +
      geom_line() +
        labs(
          x = "Days lasted",
          y = "Proportion remaining",
          colour = "Gender",
          title = "Survival curves",
          subtitle = "There is some evidence that, on average, women tend to survive longer than men",
          caption = "Siyu Zou"
        ) +
        theme(
          plot.title = element_text(size = 12, face = "bold",hjust = 0.5),
          plot.subtitle = element_text(size = 9, face = "italic" ,hjust = 0.5),
          )
```

```{r}
a = clean_survival |> 
      ggplot(aes(gender, days_lasted, color = gender, fill = gender)) +
      geom_boxplot(alpha = 0.5) +
      geom_jitter(width = 0.1, pch = 1, size = 3) +
      theme_minimal() +
      labs(
          colour = "Gender",
          fill = "Gender",
          x = "Gender",
          y = "Days lasted",
          title = "Survival boxplot",
          subtitle = "There is some evidence that, on average, women tend to survive longer than men",
          caption = "Siyu Zou"
      ) +
      theme(
          plot.title = element_text(size = 12, face = "bold",hjust = 0.5),
          plot.subtitle = element_text(size = 9, face = "italic" ,hjust = 0.5),
      )
```
### Survival time by gender
```{r}
a
```

## Column {data-width="500"}

### The survival time of people with different characteristics

```{r}
clean_survival %>% 
  arrange(desc(age)) %>% 
  head(80) %>% 
  select(season, age, gender, days_lasted) %>% 
  DT::datatable()
```

# Survival time comparence

## Column  

### Survival time of people with different age, stratified by gender
```{r}
clean_survival |>
      ggplot( aes(age, days_lasted) ) +
      geom_smooth() +
      facet_grid(~gender) +
      theme_minimal() +
      labs(
          x = "Age",
          y = "Days lasted",
          title = "Survival time, statified by gender",
          subtitle = "There is some evidence that, younger women tend to survive longer, older men tend to survival longer",
          caption = "Siyu Zou"
      ) +
      theme(
          plot.title = element_text(size = 8, face = "bold",hjust = 0.5),
          plot.subtitle = element_text(size = 6, face = "italic" ,hjust = 0.5),
      )
```

## Column

### Survival time  of people in different season
```{r}
clean_survival2 <- clean_survival |>
      group_by(season) |>
      mutate(total = sum(days_lasted) )
      
ggplot(clean_survival2, aes(season, total) ) +
      geom_col() +
      theme_minimal() +
      labs(
          x = "Season",
          y = "Days lasted",
          title = "Survival time, statified by season",
          subtitle = "There is some evidence that, middle season tend to survive longer",
          caption = "Siyu Zou"
      ) +
      theme(
          plot.title = element_text(size = 8, face = "bold",hjust = 0.5),
          plot.subtitle = element_text(size = 6, face = "italic" ,hjust = 0.5),
      )
```

# Survival by Gender

## Column {.sidebar}

```{r}
selectInput("gender", "Choose gender:", choices = list("Male"="Male", "Female"="Female"), selected = c("Male"))
```

## Column

```{r}
filtered_gender <- reactive({
  df %>% 
    filter(gender == input$gender)
})

renderPlot( {
      ggplot(filtered_gender(), aes(days_lasted, p)) +
      geom_line() +
        labs(
          x = "Days lasted",
          y = "Proportion remaining",
          title = "Survival curves of selected gender",
        ) +
        theme(
          plot.title = element_text(size = 12, face = "bold",hjust = 0.5),
          )
})
      
```


# Loadouts

## Column {.sidebar}

**Choose seasons to see the items was carried by different season.**

```{r}
selectInput("season", label = "Choose season:",
            choices = c(1, 2,3,4,5,6,7,8,9), selected = 1)
```

## Column

### Items number by seasons

```{r}
filtered_season <- reactive({
  loadouts %>%
    filter(season == input$season) %>% 
    group_by(item) %>%
    summarise(total_number = n())
})

renderPlot({
      ggplot(filtered_season(), aes(factor(item),total_number)) +
            geom_col( ) +
            labs(
                  title = "Item numbers carried by participants by seasons",
                  x = "seasons",
                  y = "number of items"
            )
})

```

### Most popular loadouts items

```{r}
pal <- c('#1B2624', '#3C413A', '#626262', '#7A7A7B', '#929293', '#ACABAB', '#D2D2D2', '#EBEBED')

txt <- "grey20"
line <- "grey80"
bg <- "white"

loadouts |>
  count(item) |>
  mutate(item = forcats::fct_reorder(item, n, max)) |>
  ggplot(aes(item, n)) +
  geom_col(fill = pal[1]) +
  coord_flip() +
  labs(
    title = "Most popular loadout items"
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = bg, colour = bg),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.25),
    plot.subtitle = element_text( lineheight = 0.3),
    axis.text.y = element_text(size = 7, hjust = 1, margin = margin(t = 10, b = 10, l = 10, r = 10)),
    axis.text.x = element_text(margin = margin(t = 10, b = 10, l = 10, r = 10)),
    axis.ticks = element_line(colour = line),
    axis.line = element_line(colour = line),
    panel.grid = element_line(colour = line, linetype = 3)
  )

```

# Analysis
## Column

### Plot 1: Kaplan-Meier survival curves
```{r}
df |> 
      ggplot(aes(days_lasted, p, colour = gender)) +
      geom_line() +
        labs(
          x = "Days lasted",
          y = "Proportion remaining",
          colour = "Gender",
          title = "Survival curves",
          subtitle = "There is some evidence that, on average, women tend to survive longer than men",
          caption = "Siyu Zou"
        ) +
        theme(
          plot.title = element_text(size = 16, face = "bold",hjust = 0.5),
          plot.subtitle = element_text(size = 11, face = "italic" ,hjust = 0.5),
          )
```     
 
### Plot 2: boxplots
```{r}
clean_survival |> 
      ggplot(aes(gender, days_lasted, color = gender, fill = gender)) +
      geom_boxplot(alpha = 0.5) +
      geom_jitter(width = 0.2, pch = 1, size = 3) +
      theme_minimal() +
      labs(
          colour = "Gender",
          fill = "Gender",
          x = "Gender",
          y = "Days lasted",
          title = "Survival boxplot",
          subtitle = "There is some evidence that, on average, women tend to survive longer than men",
          caption = "Siyu Zou"
      ) +
      theme(
          plot.title = element_text(size = 16, face = "bold",hjust = 0.5),
          plot.subtitle = element_text(size = 11, face = "italic" ,hjust = 0.5),
      )
```

## Column

### plot 3:
```{r}
clean_survival |>
      ggplot( aes(age, days_lasted) ) +
      geom_smooth() +
      facet_grid(~gender) +
      theme_minimal() +
      labs(
          x = "Age",
          y = "Days lasted",
          title = "Survival time by age",
          subtitle = "There is some evidence that, younger women tend to survive longer, older men tend to survival longer",
          caption = "Siyu Zou"
      ) +
      theme(
          plot.title = element_text(size = 16, face = "bold",hjust = 0.5),
          plot.subtitle = element_text(size = 10, face = "italic" ,hjust = 0.5),
      )
```

### Summarize
There is some evidence to suggest that women, on average, survive longer than men. The rules allow each survivalist to take 10 items with them. The minimum number of loadouts taken by survivalist is 1, the maximum number of loadouts taken by survivalist is 10, and the mean number is 5.5 items. 




